---
title: "Machine_Learning_SCRI_ROOT_SNPs"
author: "Houston Saxe"
date: "1/26/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(tidyr, 
               dplyr,
               tibble, 
               caret, 
               mlbench,
               caTools,
               mlr,
               randomForest,
               data.table,
               stringr)
```


## Machine learning
### Get SNP data
```{r}
M = fread('../SCRI_ROOT_SNPs_cleaned.csv')
```

# Read in rfe results
```{r}
rfe_res_all = readRDS('../rfe_results_NEM_2and3_year_length_2and3_year_CG_PHY.RData')
```

# Get NEM phenotype data
```{r}
NEM_pheno = fread('C:/Users/hsaxe/OneDrive/Documents/ALAB/GitHub/SCRI_ROOT_2/SCRI_Phenotype_Analysis/2_3_Year_RLN_Counts_and_tree_lengths.csv') %>% 
  mutate(CAL_Wip_ID = gsub(' ', '', CAL_Wip_ID) %>% 
           gsub('-', '_', .))

NEM_pheno %>% 
  filter(CAL_Wip_ID %like% 'MS1.36|MS1.41|MS1.56|MS1.122')

# CAL_Wip_ID, length_2Y, RLN_2Y, length_3Y, RLN_3Y
```

### 2Y.RLN Merge with SNP data
```{r}
datNEM_2YRLN = M %>% 
  left_join(dplyr::select(NEM_pheno, CAL_Wip_ID, RLN_2Y), 
            by = c('Hybrid' = names(NEM_pheno)[1])) %>% 
  relocate(RLN_2Y)

length(datNEM_2YRLN$RLN_2Y[is.na(datNEM_2YRLN$RLN_2Y) == T])
```

#### Drop NAs
```{r}
datNEM_2YRLN = datNEM_2YRLN %>% 
  drop_na()
```

#### What are the most important QTLs for 2Y.RLN?
```{r}
rfedat = datNEM_2YRLN %>%
  dplyr::select(!Hybrid) %>%
  mutate(across(.cols = everything(), as.numeric))

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

# res_NEM_2YRLN = rfe(RLN_2Y ~ ., rfedat, rfeControl = ctrl)
# 
# rfe_res_all$TwoY_RLN = res_NEM_2YRLN

res_NEM_2YRLN = rfe_res_all$TwoY_RLN
```

#### Distribution of phenotype data
```{r}
ggplot(datNEM_2YRLN, aes(RLN_2Y))+
 geom_density()
```


#### NEM ML model for predicting resistance
```{r}

## Train model based on rfe results
## Create binary classifier based on some threshold for prediction
set = datNEM_2YRLN %>% 
  dplyr::select(Hybrid, RLN_2Y, contains("31.01")) %>%
  mutate(Hybrid = factor(Hybrid), NEM = factor(ifelse(RLN_2Y <= 100,
                                                      "R",
                                                      "S"))) %>%
  na.omit() %>%
  dplyr::select(!RLN_2Y) %>%
  relocate(NEM)

## Set random number generator seed for splitting data
set.seed(1)

## Make 80/20 data split for training/testing
rows = createDataPartition(set$NEM, times = 1, p = 0.8, list = F)

train = set[rows,]

test = set[-rows,]

## Set random number generator seed for training model
set.seed(1)

mod = caret::train(as.formula(paste('NEM' ,
                                    paste(res_NEM_2YRLN$optVariables, collapse = ' + '),
                                    sep = ' ~ ')),
                   data = train)

mod

p = predict.train(mod, test)

cmat = confusionMatrix(p, test$NEM)

cmat

ML_res = data.frame(RLN_2Y = cmat$overall) %>% 
  bind_rows(data.frame(RLN_2Y = cmat$byClass)) %>% 
  slice(1:6, 8, 9) 

ML_res =  bind_cols(ML_res)

# epr = extractProb(models = list(mod = mod), testX = test)
# 
# # barplot(table(train$CG))
# 
# summary(train$CG)/length(train$CG)
# 
# p3 = predict(mod, test, type = "prob")
```

### 3Y.RLN Merge with SNP data
```{r}
datNEM_3YRLN = M %>% 
  left_join(dplyr::select(NEM_pheno, CAL_Wip_ID, RLN_3Y),
            by = c('Hybrid' = names(NEM_pheno)[1])) %>% 
  relocate(RLN_3Y)

length(datNEM_3YRLN$RLN_3Y[is.na(datNEM_3YRLN$RLN_3Y) == T])

NEM_missing = datNEM_3YRLN %>% 
  filter(is.na(RLN_3Y))

fwrite(NEM_missing %>% dplyr::select(RLN_3Y, Hybrid), '../Missing_NEM_Phenotypes.csv')
```

#### Drop NAs
```{r}
datNEM_3YRLN = datNEM_3YRLN %>% 
  drop_na()
```


```{r}
# rfe_res_all = list()
```


#### What are the most important QTLs for 3Y.RLN?
```{r}
rfedat = datNEM_3YRLN %>%
  dplyr::select(!Hybrid) %>%
  mutate(across(.cols = everything(), as.numeric))

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

# res_NEM_3YRLN = rfe(RLN_3Y ~ ., rfedat, rfeControl = ctrl)

# rfe_res_all$ThreeY_RLN = res_NEM_3YRLN

res_NEM_3YRLN = rfe_res_all$ThreeY_RLN
```

#### Distribution of phenotype data
```{r}
ggplot(datNEM_3YRLN, aes(RLN_3Y))+
 geom_density()
```


#### Distribution of phenotype data
```{r}
# set = datNEM_3YRLN %>% 
#   dplyr::select(Hybrid, RLN_3Y, contains("31.01")) %>%
#   mutate(Hybrid = factor(Hybrid), NEM = factor(ifelse(RLN_3Y <=15,
#                                                       "R",
#                                                       "S"))) %>%
#   na.omit() %>%
#   dplyr::select(!RLN_3Y) %>%
#   relocate(NEM)
# 
# set2 = set %>% 
#          dplyr::select(NEM, res_NEM_3YRLN$optVariables) %>% 
#          pivot_longer(!NEM, names_to = 'SNP')
# 
# ggplot(set2, aes(value, value, fill = NEM))+
#  geom_col()+
#   facet_wrap(NEM~SNP, ncol = 4)

```

```{r}
# glm = glm(as.formula(paste('NEM' , 
#                            paste(res_NEM_3YRLN$optVariables, collapse = ' + '),
#                            sep = ' ~ ')),
#     data = set %>% 
#          dplyr::select(NEM, res_NEM_3YRLN$optVariables),
#     family = binomial)
# 
# summary(glm)
```


#### NEM ML model for predicting resistance
```{r}

## Train model based on rfe results
## Create binary classifier based on some threshold for prediction
set = datNEM_3YRLN %>% 
  dplyr::select(Hybrid, RLN_3Y, contains("31.01")) %>%
  mutate(Hybrid = factor(Hybrid), NEM = factor(ifelse(RLN_3Y <= 68,
                                                      "R", 
                                                      "S"))) %>%
  na.omit() %>%
  dplyr::select(!RLN_3Y) %>%
  relocate(NEM)

## Set random number generator seed for splitting data
set.seed(1)

## Make 60/40 data split for training/testing
rows = createDataPartition(set$NEM, times = 1, p = 0.8, list = F)

train = set[rows,]

test = set[-rows,]

## Set random number generator seed for training model
set.seed(1)

mod = caret::train(as.formula(paste('NEM' , 
                                    paste(res_NEM_3YRLN$optVariables, collapse = ' + '),
                                    sep = ' ~ ')),
                   data = train)

mod

p = predict.train(mod, test)

cmat = confusionMatrix(p, test$NEM)

cmat

ML_res = data.frame(X3_RLN = cmat$overall) %>% 
  bind_rows(data.frame(X3_RLN = cmat$byClass)) %>% 
  slice(1:6, 8, 9) %>% 
  bind_cols(ML_res)

# epr = extractProb(models = list(mod = mod), testX = test)
# 
# # barplot(table(train$CG))
# 
# summary(train$CG)/length(train$CG)
# 
# p3 = predict(mod, test, type = "prob")
```

## Phenotype distribution without selection
```{r}
no_predictions_NEM = test %>%
  dplyr::select(Hybrid, NEM) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  left_join(NEM_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(no_predictions_NEM, aes(RLN_3Y))+
 geom_density()
```

## Phenotype distribution after selection 
```{r}
predictions_NEM = test %>%
  dplyr::select(Hybrid, NEM) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  filter(Predictions == 'R') %>% 
  left_join(NEM_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(predictions_NEM, aes(RLN_3Y))+
 geom_density()
```

## Stats of selection vs no selection
```{r}
predicted_proportions_NEM = predictions_NEM %>%
  mutate(mean_phenotype = mean(RLN_3Y)) %>% 
  group_by(NEM) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  # count(NEM) %>%
  pivot_wider(names_from = NEM, 
              values_from = c(n, mean_phenotype)) %>% 
  mutate(Percent_R = n_R/(n_R+n_S),
         Method = 'Predictions')

Proportion_stats_NEM = datNEM_3YRLN %>% 
  dplyr::select(Hybrid, RLN_3Y) %>% 
  mutate(Hybrid = factor(Hybrid), 
         NEM = factor(ifelse(RLN_3Y <= 68,
                             "R", 
                             "S")),
         mean_phenotype = mean(RLN_3Y)) %>% 
  na.omit() %>%
  group_by(NEM) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  pivot_wider(names_from = NEM, 
              values_from = c(n, mean_phenotype)) %>% 
   mutate(Percent_R = n_R/(n_R+n_S),
         Method = 'No Predictions') %>% 
  rbind(predicted_proportions_NEM)
  
Proportion_stats_NEM
```

## Fisher's exact test of predictions vs no predictions, 3.82 times more likely to get a resistant individual with predictions vs no predictions
```{r}
Proportion_stats_NEM %>% 
  arrange(desc(Method)) %>% 
  column_to_rownames(var = 'Method') %>% 
  dplyr::select(n_R, n_S) %>%
  as.matrix()

Proportion_stats_NEM %>% 
  arrange(desc(Method)) %>% 
  dplyr::select(n_R, n_S) %>%
  as.matrix() %>% 
  fisher.test()
```


### 2Y.length Merge with SNP data
```{r}
datNEM_2Ylength = M %>% 
  left_join(dplyr::select(NEM_pheno, CAL_Wip_ID, length_2Y), 
            by = c('Hybrid' = names(NEM_pheno)[1])) %>% 
  relocate(length_2Y)

length(datNEM_2Ylength$length_2Y[is.na(datNEM_2Ylength$length_2Y) == T])
```

#### Drop NAs
```{r}
datNEM_2Ylength = datNEM_2Ylength %>% 
  drop_na()
```



#### What are the most important QTLs for 2Y.length?
```{r}
rfedat = datNEM_2Ylength %>%
  dplyr::select(!Hybrid) %>%
  mutate(across(.cols = everything(), as.numeric))

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

# res_NEM_2Ylength = rfe(length_2Y ~ ., rfedat, rfeControl = ctrl)
# 
# rfe_res_all$TwoY_length = res_NEM_2Ylength

res_NEM_2Ylength = rfe_res_all$TwoY_length
```

#### Distribution of phenotype data
```{r}
datNEM_2Ylength = datNEM_2Ylength %>% 
  mutate(Mother = str_extract(Hybrid, 'MS1|AC'))

ggplot(datNEM_2Ylength, aes(length_2Y, color = Mother))+
 geom_density()
```

#### 2Ylength ML model for predicting resistance
```{r}

## Train model based on rfe results
## Create binary classifier based on some threshold for prediction
set = datNEM_2Ylength %>% 
  dplyr::select(Hybrid, length_2Y, contains("31.01")) %>%
  mutate(Hybrid = factor(Hybrid), NEM = factor(ifelse(length_2Y >= 75,
                                                      "Tall",
                                                      "Short"))) %>%
  na.omit() %>%
  dplyr::select(!length_2Y) %>%
  relocate(NEM)

## Set random number generator seed for splitting data
set.seed(1)

## Make 60/40 data split for training/testing
rows = createDataPartition(set$NEM, times = 1, p =0.8, list = F)

train = set[rows,]

test = set[-rows,]

## Set random number generator seed for training model
set.seed(1)

mod = caret::train(as.formula(paste('NEM' , 
                                    paste(res_NEM_2Ylength$optVariables, collapse = ' + '),
                                    sep = ' ~ ')),
                   data = train)

mod

p = predict.train(mod, test)

cmat = confusionMatrix(p, test$NEM)

cmat

ML_res = data.frame(length_2Y = cmat$overall) %>% 
  bind_rows(data.frame(length_2Y = cmat$byClass)) %>% 
  slice(1:6, 8, 9) %>% 
  bind_cols(ML_res)

# epr = extractProb(models = list(mod = mod), testX = test)
# 
# # barplot(table(train$CG))
# 
# summary(train$CG)/length(train$CG)
# 
# p3 = predict(mod, test, type = "prob")
```

### 3Y.length Merge with SNP data
```{r}
datNEM_3Ylength = M %>% 
  left_join(dplyr::select(NEM_pheno, CAL_Wip_ID, length_3Y),
            by = c('Hybrid' = names(NEM_pheno)[1])) %>% 
  relocate(length_3Y)

length(datNEM_3Ylength$length_3Y[is.na(datNEM_3Ylength$length_3Y) == T])
```

#### Drop NAs
```{r}
datNEM_3Ylength = datNEM_3Ylength %>% 
  drop_na()
```

#### What are the most important QTLs for 3Y.length?
```{r}
rfedat = datNEM_3Ylength %>%
  dplyr::select(!Hybrid) %>%
  mutate(across(.cols = everything(), as.numeric))

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

# res_NEM_3Ylength = rfe(length_3Y ~ ., rfedat, rfeControl = ctrl)
# 
# rfe_res_all$ThreeY_length = res_NEM_3Ylength

res_NEM_3Ylength = rfe_res_all$ThreeY_length

res_NEM_3Ylength$optVariables = res_NEM_3Ylength$optVariables[1:5]
```

#### Distribution of phenotype data
```{r}
ggplot(datNEM_3Ylength, aes(length_3Y))+
 geom_density()
```

#### Length_3Y ML model for predicting resistance
```{r}

## Train model based on rfe results
## Create binary classifier based on some threshold for prediction
set = datNEM_3Ylength %>% 
  dplyr::select(Hybrid, length_3Y, contains("31.01")) %>%
  mutate(Hybrid = factor(Hybrid),
         Length_3Y = factor(ifelse(length_3Y >= 100, 
                                                      "Tall",
                                                      "Short"))) %>%
  na.omit() %>%
  dplyr::select(!length_3Y) %>%
  relocate(Length_3Y)

## Set random number generator seed for splitting data
set.seed(1)

## Make 60/40 data split for training/testing
rows = createDataPartition(set$Length_3Y, times = 1, p = 0.8, list = F)

train = set[rows,]

test = set[-rows,]

## Set random number generator seed for training model
set.seed(1)

mod = caret::train(as.formula(paste('Length_3Y' ,
                                    paste(res_NEM_3Ylength$optVariables, collapse = ' + '),
                                    sep = ' ~ ')),
                   data = train)

mod

p = predict.train(mod, test)

cmat = confusionMatrix(p, test$Length_3Y)

cmat

ML_res = data.frame(length_3Y = cmat$overall) %>% 
  bind_rows(data.frame(length_3Y = cmat$byClass)) %>% 
  slice(1:6, 8, 9) %>% 
  bind_cols(ML_res)

# epr = extractProb(models = list(mod = mod), testX = test)
# 
# # barplot(table(train$CG))
# 
# summary(train$CG)/length(train$CG)
# 

```


```{r}
# library(verification)

p3 = predict.train(mod, test, type = "prob")

p3

df2 = data.frame(probs = p3, 
                 labels = test$Length_3Y,
                 predictions = p) %>% 
  mutate(predictions_N = ifelse(predictions == 'Short', 1, 0),
         labels_N = ifelse(labels == 'Short', 1, 0))

# library(ROCR)
# 
# pred2 = prediction(df2$probs.Short, df2$labels_N)
# 
# perf2 = performance(pred2, 'tpr', 'fpr')
# 
# plot(perf2, colorize = T)
```

```{r}
# library(plotROC)
# 
# ggplot(df2, aes(m = probs.Short, d = labels_N))+
#   geom_roc(n.cuts = 20, labels = T)
```


## Phenotype distribution without selection
```{r}
no_predictions_Length_3Y = test %>%
  dplyr::select(Hybrid, Length_3Y) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  left_join(NEM_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(no_predictions_Length_3Y, aes(length_3Y))+
 geom_density()
```

## Phenotype distribution after selection 
```{r}
predictions_Length_3Y = test %>%
  dplyr::select(Hybrid, Length_3Y) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  filter(Predictions == 'Tall') %>% 
  left_join(NEM_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(predictions_Length_3Y, aes(length_3Y))+
 geom_density()
```

## Stats of selection vs no selection
```{r}
predicted_proportions_Length_3Y = predictions_Length_3Y %>%
  mutate(mean_phenotype = mean(length_3Y)) %>% 
  group_by(Length_3Y) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  # count(Length_3Y) %>%
  pivot_wider(names_from = Length_3Y, 
              values_from = c(n, mean_phenotype)) %>% 
  mutate(Percent_Tall = n_Tall/(n_Tall+n_Short),
         Method = 'Predictions')

Proportion_stats_Length_3Y = datNEM_3Ylength %>% 
  dplyr::select(Hybrid, length_3Y) %>% 
  mutate(Hybrid = factor(Hybrid), 
         Length_3Y = factor(ifelse(length_3Y >= 100, 
                                                      "Tall",
                                                      "Short")),
         mean_phenotype = mean(length_3Y)) %>% 
  na.omit() %>%
  group_by(Length_3Y) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  pivot_wider(names_from = Length_3Y, 
              values_from = c(n, mean_phenotype)) %>% 
   mutate(Percent_Tall = n_Tall/(n_Tall+n_Short),
         Method = 'No Predictions') %>% 
  rbind(predicted_proportions_Length_3Y)
  
Proportion_stats_Length_3Y  
```

## Fisher's exact test of predictions vs no predictions, 2.48 times more likely to get a tall individual with predictions vs no predictions
```{r}
Proportion_stats_Length_3Y %>% 
  arrange(desc(Method)) %>% 
  column_to_rownames(var = 'Method') %>% 
  dplyr::select(n_Tall, n_Short) %>%
  as.matrix()

Proportion_stats_Length_3Y %>% 
  arrange(desc(Method)) %>% 
  dplyr::select(n_Tall, n_Short) %>%
  as.matrix() %>% 
  fisher.test()
```

# Get CG phenotype data
```{r}
## What are the most important variables for this model?
CG_pheno = fread('C:/Users/hsaxe/OneDrive/Documents/ALAB/GitHub/SCRI_ROOT_2/SCRI_Phenotype_Analysis/CG_Avg_Score.csv') %>% 
  mutate(CAL_Wip_ID = gsub(' ', '', CAL_Wip_ID) %>% 
           gsub('-', '_', .))
```

#### Merge with SNP data
```{r}
datCG = M %>% 
  left_join(CG_pheno, by = c('Hybrid' = names(CG_pheno)[1])) %>% 
  relocate(CG_Avg)

length(datCG$CG_Avg[is.na(datCG$CG_Avg) == T])
```

#### Drop NAs
```{r}
datCG = datCG %>% 
  drop_na()

dim(datCG)
```

#### What are the most important QTLs for CG?
```{r}
rfedat = datCG %>%
  dplyr::select(!Hybrid) %>%
  mutate(across(.cols = everything(), as.numeric))

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

# res_CG = rfe(CG_Avg ~ ., rfedat, rfeControl = ctrl)
# 
# rfe_res_all$CG = res_CG 

res_CG = rfe_res_all$CG
```


#### Distribution of phenotype data
```{r}
ggplot(datCG, aes(CG_Avg))+
 geom_density()
```

#### CG ML model for predicting resistance
```{r}

## Train model based on rfe results
## Create binary classifier based on some threshold for prediction
set = datCG %>% dplyr::select(Hybrid, CG_Avg, contains("31.01")) %>% 
  mutate(Hybrid = factor(Hybrid), CG = factor(ifelse(CG_Avg <= 2,
                                                     "R", 
                                                     "S"))) %>% 
  na.omit() %>% 
  dplyr::select(!CG_Avg) %>%
  relocate(CG)

## Set random number generator seed for splitting data
set.seed(1)

## Make 60/40 data split for training/testing
rows = createDataPartition(set$CG, times = 1, p = 0.6, list = F)

train = set[rows,]

test = set[-rows,]

## Set random number generator seed for training model
set.seed(1)

mod = caret::train(as.formula(paste('CG' , 
                                    paste(res_CG$optVariables, collapse = ' + '),
                                    sep = ' ~ ')),
                   data = train)

mod

p = predict.train(mod, test)

cmat = confusionMatrix(p, test$CG)

cmat

ML_res = data.frame(CG = cmat$overall) %>% 
  bind_rows(data.frame(CG = cmat$byClass)) %>% 
  slice(1:6, 8, 9) %>% 
  bind_cols(ML_res)

# epr = extractProb(models = list(mod = mod), testX = test)
# 
# # barplot(table(train$CG))
# 
# summary(train$CG)/length(train$CG)
# 
# p3 = predict(mod, test, type = "prob")
```

## Phenotype distribution without selection
```{r}
no_predictions_CG = test %>%
  dplyr::select(Hybrid, CG) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  left_join(CG_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(no_predictions_CG, aes(CG_Avg))+
 geom_density()
```

## Phenotype distribution after selection 
```{r}
predictions_CG = test %>%
  dplyr::select(Hybrid, CG) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  filter(Predictions == 'R') %>% 
  left_join(CG_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(predictions_CG, aes(CG_Avg))+
 geom_density()
```

## Stats of selection vs no selection
```{r}
predicted_proportions_CG = predictions_CG %>%
  mutate(mean_phenotype = mean(CG_Avg)) %>% 
  group_by(CG) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  # count(CG) %>%
  pivot_wider(names_from = CG, 
              values_from = c(n, mean_phenotype)) %>% 
  mutate(Percent_R = n_R/(n_R+n_S),
         Method = 'Predictions')

Proportion_stats_CG = datCG %>% 
  dplyr::select(Hybrid, CG_Avg) %>% 
  mutate(Hybrid = factor(Hybrid), 
         CG = factor(ifelse(CG_Avg <= 2,
                             "R", 
                             "S")),
         mean_phenotype = mean(CG_Avg)) %>% 
  na.omit() %>%
  group_by(CG) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  pivot_wider(names_from = CG, 
              values_from = c(n, mean_phenotype)) %>% 
   mutate(Percent_R = n_R/(n_R+n_S),
         Method = 'No Predictions') %>% 
  rbind(predicted_proportions_CG)
  
Proportion_stats_CG  
```

## Fisher's exact test of predictions vs no predictions, 22.74 times more likely to get a resistant individual with predictions vs no predictions
```{r}
Proportion_stats_CG %>% 
  arrange(desc(Method)) %>% 
  column_to_rownames(var = 'Method') %>% 
  dplyr::select(n_R, n_S) %>%
  as.matrix()

Proportion_stats_CG %>% 
  arrange(desc(Method)) %>% 
  dplyr::select(n_R, n_S) %>%
  as.matrix() %>% 
  fisher.test()
```


# Get PHY phenotype data
```{r}
## What are the most important variables for this model?
PHY_pheno = fread('C:/Users/hsaxe/OneDrive/Documents/ALAB/GitHub/SCRI_ROOT_2/SCRI_Phenotype_Analysis/PHY_Avg_Score.csv') %>% 
  mutate(CAL_Wip_ID = gsub(' ', '', CAL_Wip_ID) %>% 
           gsub('-', '_', .))
```

#### Merge with SNP data
```{r}
datPHY = M %>% 
  left_join(PHY_pheno, by = c('Hybrid' = names(PHY_pheno)[1])) %>% 
  relocate(PHY_Avg)

length(datPHY$PHY_Avg[is.na(datPHY$PHY_Avg) == T])
```

#### Drop NAs
```{r}
datPHY = datPHY %>% 
  drop_na()
```


### What are the most important QTLs for PHY?
```{r}
rfedat = datPHY %>%
  dplyr::select(!Hybrid) %>%
  mutate(across(.cols = everything(), as.numeric))

ctrl <- rfeControl(functions = rfFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   verbose = FALSE)

# res_PHY = rfe(PHY_Avg ~ ., rfedat, rfeControl = ctrl)
# 
# rfe_res_all$PHY = res_PHY 


res_PHY = rfe_res_all$PHY
```

#### Distribution of phenotype data
```{r}
ggplot(datPHY, aes(PHY_Avg))+
 geom_density()
```

#### PHY ML model for predicting resistance
```{r}

## Train model based on rfe results
## Create binary classifier based on some threshold for prediction
set = datPHY %>% dplyr::select(Hybrid, PHY_Avg, contains("31.01")) %>% 
  mutate(Hybrid = factor(Hybrid), PHY = factor(ifelse(PHY_Avg <= 25,
                                                      "R", 
                                                      "S"))) %>% 
  na.omit() %>% 
  dplyr::select(!PHY_Avg) %>%
  relocate(PHY)

## Set random number generator seed for splitting data
set.seed(1)

## Make 80/20 data split for training/testing
rows = createDataPartition(set$PHY, times = 1, p = 0.8, list = F)

train = set[rows,]

test = set[-rows,]

## Set random number generator seed for training model
set.seed(1)

mod = caret::train(as.formula(paste('PHY' , 
                                    paste(res_PHY$optVariables, collapse = ' + '), 
                                    sep = ' ~ ')),
                   data = train)

mod

p = predict.train(mod, test)

cmat = confusionMatrix(p, test$PHY)

cmat

ML_res = data.frame(PHY = cmat$overall) %>% 
  bind_rows(data.frame(PHY = cmat$byClass)) %>% 
  slice(1:6, 8, 9) %>% 
  bind_cols(ML_res)

predictions_PHY = test %>%
  dplyr::select(Hybrid, PHY) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  filter(Predictions == 'R') %>% 
  left_join(PHY_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))



# epr = extractProb(models = list(mod = mod), testX = test)
# 
# # barplot(table(train$PHY))
# 
# summary(train$PHY)/length(train$PHY)
# 
# p3 = predict(mod, test, type = "prob")
```

## Phenotype distribution without selection
```{r}
no_predictions_PHY = test %>%
  dplyr::select(Hybrid, PHY) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  left_join(PHY_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(no_predictions_PHY, aes(PHY_Avg))+
 geom_density()
```

## Phenotype distribution after selection 
```{r}
predictions_PHY = test %>%
  dplyr::select(Hybrid, PHY) %>% 
  mutate(Predictions = p) %>% 
  relocate(Predictions) %>% 
  filter(Predictions == 'R') %>% 
  left_join(PHY_pheno, by = c('Hybrid' = 'CAL_Wip_ID'))

ggplot(predictions_PHY, aes(PHY_Avg))+
 geom_density()
```
## Stats of selection vs no selection
```{r}
predicted_proportions_PHY = predictions_PHY %>%
  mutate(mean_phenotype = mean(PHY_Avg)) %>% 
  group_by(PHY) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  # count(PHY) %>%
  pivot_wider(names_from = PHY, 
              values_from = c(n, mean_phenotype)) %>% 
  mutate(Percent_R = n_R/(n_R+n_S),
         Method = 'Predictions')

Proportion_stats_PHY = datPHY %>% 
  dplyr::select(Hybrid, PHY_Avg) %>% 
  mutate(Hybrid = factor(Hybrid), 
         PHY = factor(ifelse(PHY_Avg <= 25,
                             "R", 
                             "S")),
         mean_phenotype = mean(PHY_Avg)) %>% 
  na.omit() %>%
  group_by(PHY) %>%
  summarise(n = n(),
            mean_phenotype = mean_phenotype) %>% 
  distinct() %>% 
  pivot_wider(names_from = PHY, 
              values_from = c(n, mean_phenotype)) %>% 
   mutate(Percent_R = n_R/(n_R+n_S),
         Method = 'No Predictions') %>% 
  rbind(predicted_proportions_PHY)
  
Proportion_stats_PHY  
```

## Fisher's exact test of predictions vs no predictions, 5.26 times more likely to get a resistant individual with predictions vs no predictions
```{r}
Proportion_stats_PHY %>% 
  arrange(desc(Method)) %>% 
  column_to_rownames(var = 'Method') %>% 
  dplyr::select(n_R, n_S) %>%
  as.matrix()

Proportion_stats_PHY %>% 
  arrange(desc(Method)) %>% 
  dplyr::select(n_R, n_S) %>%
  as.matrix() %>% 
  fisher.test()
```

```{r}
# ML_resF = ML_res %>% 
#   dplyr::select( RLN_2Y, RLN_3Y, length_2Y, length_3Y, CG, PHY) %>% 
#   rownames_to_column(var = 'Statistic') %>% 
#   add_row(Statistic = "Threshold", RLN_2Y = 15, RLN_3Y = 100, length_2Y = 75, length_3Y = 75, CG = 2, PHY = 25) %>% 
#   mutate(across(2:7, ~ round(.x, 3)))
# 
# fwrite(ML_resF, '../ML_res_all.csv')
```


# Saving rfe results
```{r}
# rfe_res_all = list(TwoY_RLN = res_NEM_2YRLN,
#                    ThreeY_RLN = res_NEM_3YRLN,
#                    TwoY_length = res_NEM_2Ylength,
#                    ThreeY_length = res_NEM_3Ylength,
#                    CG = res_CG,
#                    PHY = res_PHY)
# 
# rfe_res_all
# 
# saveRDS(rfe_res_all, '../rfe_results_NEM_2and3_year_length_2and3_year_CG_PHY.RData')
```

```{r}
# View(rfe_res_all)
# 
# opt_vars = sapply(rfe_res_all, FUN = function(x){
#  
#   a = data.frame(seq(1:4)) 
#   
#   a = cbind(a, head(x$optVariables, n = 4))
#   
# })
# 
# opt_vars = apply(opt_vars, MARGIN = 2, FUN = function(x){
#  
#   unlist(x)
#   
# })
# 
# opt_vars = opt_vars %>%
#   as.data.frame() %>% 
#   slice(5:8)
# 
# fwrite(opt_vars, 'opt_vars_CG_PHY_NEM_LENGTH.csv')

```
